---
title: "ECMA 31360, PSet 2: Solutions"
author: "YOUR NAMES, University of Chicago"
geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
date: "XX-XX-2026"
output: pdf_document
header-includes:
  - \usepackage{caption}
  - \usepackage{float}
  - \usepackage{xcolor}
  - \usepackage{amsmath}
  - \usepackage{enumitem}
---

```{r setup-repos, include=FALSE}
# Ensure CRAN is set for non-interactive knit sessions
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H", out.extra = "")
```


# Carry-over from PSet2 

<!-- Insert here any script from PSet2 that you need for PSet3 -->

---

# (\textcolor{blue}{[ \ \ ]} out of 50p) PART I: Test Balance in the Observed Predetermined Variables (OPVs)

## (\textcolor{blue}{[ \ \ ]} out of 15p) Q1: Read and Understand the NSW Application's Companion Document

<!-- Just write DONE -->

---

## (\textcolor{blue}{[ \ \ ]} out of 23p) Q2: Implement Procedure 4 (SUR Estimation followed by Joint Testing)

### (\textcolor{blue}{[ \ \ ]} out of 10p) Q2.a: SUR Estimation

### Script and Output

---

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q2.b: Compare FGLS to OLS equation-by-equation

---

### (\textcolor{blue}{[ \ \ ]} out of 3p) Q2.c: Take a closer look at the variance-covariance matrix

### Script and Output

### Commentary

---

### (\textcolor{blue}{[ \ \ ]} out of 4p) Q2.d: Implement Joint Test Manually

### Script and Output

---

### (\textcolor{blue}{[ \ \ ]} out of 4p) Q2.e: Implement Joint Test Automatically

### Script and Output

### Commentary

---

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q2.f: Decision

---


## (\textcolor{blue}{[ \ \ ]} out of 3p) Q3: Implement Procedure 5 (Hotelling's T2 Test)

### Script and Output

### Commentary

---

## (\textcolor{blue}{[ \ \ ]} out of 4p) Q4: Implement Procedure 6 (OPV do not predict treatment assignment)

### Script and Output

### Commentary

---

## (\textcolor{blue}{[ \ \ ]} out of 5p) Q5: Implement Procedures 7 and 8 (control for FWER)

### Script and Output

### Commentary

---

# (\textcolor{blue}{[ \ \ ]} out of 50p) PART II: Review of OLS for Causal Analysis

## (\textcolor{blue}{[ \ \ ]} out of 3p) Q6: Two ways of obtaining the DM estimator

### Script and Output

### Commentary

---

## (\textcolor{blue}{[ \ \ ]} out of 4p) Q7: Fully-saturated specification in Nodegree

### Script and Output

---

## (\textcolor{blue}{[ \ \ ]} out of 2p) Q8: Get $\widehat{ATE}$ in one step

### Script and Output

---

## (\textcolor{blue}{[ \ \ ]} out of 22p) Q9: Implications of estimating a not-fully saturated specification in an OPV that takes $M$ distinct values

---

## (\textcolor{blue}{[ \ \ ]} out of 14p) Q10: Implement the DM estimator with Regression Adjustment (various specifications)

### (\textcolor{blue}{[ \ \ ]} out of 3p) Q10.a: Estimate Specifications 2, 3 and 4

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q10.a.i: Estimate Specification 2

### Script and Output

### Commentary

---

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q10.a.ii: Estimate Specification 3

### Script and Output

### Commentary

---

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q10.a.iii: Estimate Specification 4

### Script and Output

### Commentary

---

### (\textcolor{blue}{[ \ \ ]} out of 5p) Q10.b: Reasons to include OPVs when they are balanced

---

### (\textcolor{blue}{[ \ \ ]} out of 1p) Q10.c:  Is it problematic to regression-adjust for OPVs that are lagged outcomes?

---

### (\textcolor{blue}{[ \ \ ]} out of 5p) Q10.d:  Interactions of OPV with Treatment Indicator and Two Hypothesis Testing Problems

### Script and Output

### Commentary

---


## (\textcolor{blue}{[ \ \ ]} out of 5p) Q11: Mechanisms for NSW intervention to impact post-intervention earnings

